---
title: "Chapter 11: Probability models for counts"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
library(pdftools)
library(tidyverse)
```

# Learning outcomes

By the end of this lesson you should:

-   Know how to use the binomial distribution, as well as its mean, and variance.
-   Know how to use the Poisson distribution, as well as its mean, and variance.
-   Understand differences between the binomial and Poisson distributions, and when the two could be applied.

# Bernoulli trial

::: notes
find the E(B) and Var(B)
:::

::: incremental
-   flipping a coin

-   insured or not insured

-   luggage lost or not lost

-   $$\begin{equation}
    B =\begin{cases}
    1, & \text{if the trial is a success}.\\
    0, & \text{if the trial is not a success}.
    \end{cases}
    \end{equation}$$

-   $p$ is the probability of success
:::

## E(Bernoulli)

::: incremental
-   $E(B) = 0 \cdot P(B=0) + 1 \cdot P(B=1)$
-   $$\begin{aligned}
    E(B) &= 0 \cdot P(B=0) + 1 \cdot P(B=1)\\
    &= 0 \cdot (1-p) + 1 \cdot p\\
    &= p
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 2, seconds = 0)
```

## Var(Bernoulli)

::: notes
We usually don't want this information about 1 person, we want it about a whole plane full of people, or a whole town.

In that case, we sum together $n$ independent Bernoulli trials.
:::

::: incremental
-   $Var(B) = (0 - p)^2 \cdot P(B=0) + (1 - p)^2 \cdot P(B=1)$
-   $$\begin{aligned}
    Var(B) &= (0 - p)^2 \cdot P(B=0) + (1 - p)^2 \cdot P(B=1)\\
    &= p^2 \cdot (1-p) + (1 - p)^2 \cdot p\\
    &= p^2 - p^3 + p(1 - 2p + p^2) \\
    &= p^2 - p^3 + p - 2p^2 + p^3\\
    &= p(1-p)
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 2, seconds = 0)
```

## Multiple Bernoulli trials

::: notest
This happens so often that we have another name for this distribution
:::

$$Y = B_1 + B_2 + ...  + B_n$$

::: incremental
-   $Y$ is the sum of independent and identically distributed random variables
-   What is the [mean and variance](https://stats1010-f22.github.io/website/slides/lect_09.html#/addition-rules-for-iid-variables)?
:::

# Binomial distribution - FIST

::: notes
The order of successes?
:::

The number of successes in $n$ trials

::: incremental
-   Binomial - Biceps - FIST

-   F - fixed number of trials

-   I - independent outcomes

-   S - probability of success is constant

-   T - two outcomes
:::

## E(Binomial)

$$E(Y) = E(B_1 + B_2 + ... + B_n)$$

::: incremental
-   $$\begin{aligned}
    E(Y) &= E(B_1) + E(B_2) + ... + E(B_n) \\
    &= p + p + ... + p \\
    &= np
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 2, seconds = 0)
```

## Var(Binomial) {.smaller}

::: notes
there are multiple ways that we can have 5 successes, so we must also learn to count
:::

$$Var(Y) = Var(B_1 + B_2 + ... + B_n)$$

::: incremental
-   $$\begin{aligned}
    Var(Y) &= Var(B_1) + Var(B_2) + ... + Var(B_n) \\
    &= p(1-p) + p(1-p) + ... + p(1-p) \\
    &= np(1-p)
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 2, seconds = 0)
```

## Counting things {.smaller}

::: notes
we just went into groups how many groups are possible?
:::

In a class of 85 students how many groups of 4 are possible?

::: incremental
-   $85 \cdot 84 \cdot 83 \cdot 82$

-   $48,594,840$ groups

-   If we don't care about order

-   $\frac{85 \cdot 84 \cdot 83 \cdot 82}{4 \cdot 3 \cdot 2 \cdot 1}$

-   $2,024,785$ groups

-   *Combination* written as ${}_{n}C_{k}$ here $n = 85$, $k = 4$ or $\binom{85}{4}$
:::

```{r}
#| echo: true
#| eval: true

exp(lfactorial(85))/exp(lfactorial(81))


exp(lfactorial(85))/(exp(lfactorial(81))*exp(lfactorial(4)))
```

## Binomial pdf

If $Y \sim Bin(n, p)$ where

-   $n = $ number of trials
-   $p = $ probability of success
-   $y = $ number of successes in $n$ Bernoulli trials

<br>

then <br> $$P(Y = y) = \binom{n}{y}p^y(1-p)^{n-y}$$

## Example

::: incremental
1.  Let $Y \sim Bin(n = 5, p = 0.2)$ find the $E(Y)$ and $Var(Y)$

```{=html}
<!-- -->
```
a.  $E(Y) = np = 1$
b.  $Var(Y) = np(1-p) = 0.8$
c.  What is the probability that $y = 3$? In `R: dbinom(size  = 5, prob = 0.2, x = 3)` $0.0512$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 2, seconds = 0)
```

## Warnings

10% Condition: if trials are selected at random, it is OK to ignore dependence caused by sampling from a finite population if the selected trials make up less than 10% of the population

## Limit - mins

::: notes
some of these cars may come by in the same minute..
:::

I know about 7 cars drive by my house in an hour, what's the probability 18 cars drive by in the next 60 mins?

::: incremental
-   $p = \frac{7}{60}$

-   $Y \sim Bin(60, \frac{7}{60})$

-   $\binom{60}{18}p^{18}(1-p)^{42}$
:::

## Limit - secs

I know about 7 cars drive by my house in an hour, what's the probability 18 cars drive by in the next 3600 secs?

::: incremental
-   $p = \frac{7}{3600}$

-   $Y \sim Bin(3600, \frac{7}{3600})$

-   $\binom{3600}{18}p^{18}(1-p)^{3582}$
:::

## Limit - smallest interval

I know about 7 cars drive by my house in an hour, what's the probability 18 cars drive by in the next much smaller than a second? Full derivation can be found [here](https://stats1010-f22.github.io/website/supplemental/chi-squared.html).

::: columns
::: {.column width="50%"}
::: incremental
-   $\binom{n}{18}(\frac{7}{n})^{18}(1-\frac{7}{n})^{n-18}$

-   $\frac{n\cdot(n-1)\ldots(n-17)}{18!}\cdot \frac{7^{18}}{n^{18}}\cdot (1-\frac{7}{n})^{n} \cdot \frac{1}{(1-\frac{7}{n})^{18}}$

-   $\frac{n\cdot(n-1)\ldots(n-17)}{n^{18}} \rightarrow 1$

-   $\frac{1}{(1-\frac{7}{n})^{18}} \rightarrow 1$
:::
:::

::: {.column width="50%"}
::: incremental
$\frac{7^{18}}{18!}e^{-7}$
:::
:::
:::

# Poisson distribution

If $X \sim Pois(\lambda)$ then $P(X=x) = \frac{\lambda^{x}}{x!}e^{-\lambda}$

## Poisson distribution - RIPS

::: incremental
-   Poisson - [fish](http://pecheur.info/en/gollum-fish-song.html) - Rips
-   R - randomly through space or time
-   I - indepedent
-   P - proportional to interval size
-   S - singly - no multiple occurences in space or time
:::

## E(Pois)

::: incremental
$\lambda$
:::

## Var(Pois)

::: incremental
$\lambda$
:::

## Example

::: incremental
1.  Let $Y \sim Pois(\lambda = 5)$ find the $E(Y)$ and $Var(Y)$

```{=html}
<!-- -->
```
a.  $E(Y) = 5$
b.  $Var(Y) = 5$
c.  What is the probability that $y = 3$? In `R: dpois(x = 3, lambda = 5)` $0.1403739$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 2, seconds = 0)
```
