---
title: "Lec 3 - Exploring categorical data"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
```

# Learning outcomes

By the end of this lesson you should:

-   Be able to identify categorical variables and why they are important

-   Graphical representation of two categorical variables in R

-   Tabular representation of two categorical variables

-   Graphical representation of one categorical variable

# Word cloud

Use this [link](PollEv.com/gwynnsturdevant155), or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("PollEv.com/gwynnsturdevant155")
plot(qr)
```

## Definition of categorical variable

A categorical or qualitative variable is a variable that can not be counted. They are descriptors that are or grouping factors.

## The purpose of exploring categorical variables

-   Exploratory Data Analysis is about learning the structure of a dataset through a series of numerical and graphical techniques.

-   When you do EDA, you'll look for both

    -   general trends and

    -   interesting outliers in your data.

generate questions that will help inform subsequent analysis.

# Counts & proportions

-   How many freshman are there in our class?

-   What proportion of our class turned in the diamonds assignment?

-   How many transportation stocks are in our portfolio?

-   DOES NOT RESTRICT THE POPULATION

-   Others?

    # Marginal distribution

-   The count of the distribution of one variable.

-   This is referred to asmthe marginal distribution because in a contingency table, we usually compute the column sums and row sums in the **margins**.

    # Conditional probabilities

-   Of all the freshman students in our class what percent turned in the diamonds exercise?

-   Of all the students who turned in the diamonds exercise how many were freshman?

-   How many transportation stocks are in our portfolio that performed well last year?

-   Of all the stocks in our portfolio that performed well last year, how many are transportation stocks?

-   RESTRICTED TO A SUBPOPULATION

Others?

# Words to watch for

-   Mutually exclusive

-   Associated

-   Independence

-   Chi-Squared test

-   $H_0$ and $H_A$

-   Contingency table

# Application exercise

[Click here](https://stats1010-f22.github.io/website/ae/ae-2-exploring-categorical-vars.html) or the qr code below to write your first line of code

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-2-exploring-categorical-vars.html")
plot(qr)
```
