---
title: "Lec 5 - Introduction to probability"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
```

# Learning outcomes

By the end of this lesson you should:

-   Understand probability and its axioms

-   Know and understand words like union, intersection, and compliment

-   Be able to calculate basic probabilities

# History of probability

-   Gives you an edge in games of chance
-   Cardano, Fermat, Pascal, Kolmogorov
-   casinos rely on probability theory for profits

# Word cloud of probability

Use this [link](PollEv.com/gwynnsturdevant155), or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("PollEv.com/gwynnsturdevant155")
plot(qr)
```

# Basic terminology

-   An *experiment* is a process that produces an *observation*.
-   An *outcome* is a possible observation.
-   The set of all possible outcomes is called the *sample space*.
-   An *event* is a subset of the sample space.
-   A *trial* is a single running of an experiment.
-   Events are *disjoint* or *mutually exclusiv*e if they have no outcomes in common.

# An *experiment*

Rolling a fair, 6-sided and observing the number of pips that appear on top.

1.  What is one *outcome* of a *trial*?
2.  What is the *sample space*?
3.  What are possible *events*?
4.  What are some *disjoint* events?
5.  What are the *probabilities* of the events above?

# An *experiment*

```{r}
countdown(minutes = 5, seconds = 0, style = "position: relative; width: min-content;")
```

# The 3 axioms of probability

1.  The probability of the *sample space* ($S$) is 1 (ie $P(S) = 1$)
2.  A probability $p$ is always between $0$ and $1$ (ie $0 \leq p \leq 1$)
3.  If two events $A$ and $B$ are *disjoint*, then $P(A \cup B) = P(A) + P(B)$

# One event {background-image="lec_05-images/one_event.png" background-size="contain"}



# Your turn

[Click here](https://stats1010-f22.github.io/website/ae/ae-3.html) or the qr code below to write your first line of code

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-3.htm")
plot(qr)
```
