---
title: "Chapter 11: Probability models for counts"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
library(pdftools)
library(tidyverse)
```

# Learning outcomes

By the end of this lesson you should:

-   Know how to use the binomial distribution, as well as its mean, and variance.
-   Know how to use the Poisson distribution, as well as its mean, and variance.
-   Understand differences between the binomial and Poisson distributions, and when the two could be applied.

# Bernoulli trial

::: notes
find the E(B) and Var(B)
:::

::: incremental
-   flipping a coin

-   insured or not insured

-   luggage lost or not lost

-   $$\begin{equation}
    B =\begin{cases}
    1, & \text{if the trial is a success}.\\
    0, & \text{if the trial is a success}.
    \end{cases}
    \end{equation}$$

-   $p$ is the probability of success
:::

## E(Bernoulli)

::: incremental
-   $E(B) = 0 \cdot P(B=0) + 1 \cdot P(B=1)$
-   $$\begin{aligned}
    E(B) &= 0 \cdot P(B=0) + 1 \cdot P(B=1)\\
    &= 0 \cdot (1-p) + 1 \cdot p\\
    &= p
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 3, seconds = 0)
```

## Var(Bernoulli)

::: notes
We usually don't want this information about 1 person, we want it about a whole plane full of people, or a whole town.

In that case, we sum together $n$ independent Bernoulli trials.
:::

::: incremental
-   $Var(B) = (0 - p)^2 \cdot P(B=0) + (1 - p)^2 \cdot P(B=1)$
-   $$\begin{aligned}
    Var(B) &= (0 - p)^2 \cdot P(B=0) + (1 - p)^2 \cdot P(B=1)\\
    &= p^2 \cdot (1-p) + (1 - p)^2 \cdot p\\
    &= p^2 - p^3 + p(1 - 2p + p^2) \\
    &= p^2 - p^3 + p - 2p^2 + p^3\\
    &= p(1-p)
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 3, seconds = 0)
```

## Multiple Bernoulli trials

::: notest
This happens so often that we have another name for this distribution
:::

$$Y = B_1 + B_2 + B_3 + B_4 + B_5 + B_6 + B_7$$

::: incremental
-   $Y$ is the sum of independent and identically distributed random variables
-   What is the [mean and variance](https://stats1010-f22.github.io/website/slides/lect_09.html#/addition-rules-for-iid-variables)?
:::

# Binomial distribution - FIST

::: notes
The order of successes?
:::

The number of successes in $n$ trials

::: incremental
-   Binomial - Biceps - FIST

-   F - fixed number of trials

-   I - independent outcomes

-   S - probability of success is constant

-   T - two outcomes
:::

## E(Binomial)

$$E(Y) = E(B_1 + B_2 + ... + B_n)$$

::: incremental
-   $$\begin{aligned}
    E(Y) &= E(B_1) + E(B_2) + ... + E(B_n) \\
    &= p + p + ... + p \\
    &= np
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 3, seconds = 0)
```

## Var(Binomial) {.smaller}

::: notes
there are multiple ways that we can have 5 successes, so we must also learn to count
:::

$$Var(Y) = Var(B_1 + B_2 + ... + B_n)$$

::: incremental
-   $$\begin{aligned}
    Var(Y) &= Var(B_1) + Var(B_2) + ... + Var(B_n) \\
    &= p(1-p) + p(1-p) + ... + p(1-p) \\
    &= np(1-p)
    \end{aligned}$$
:::

```{r}
#| echo: false
#| eval: true

countdown(minutes = 3, seconds = 0)
```

## Counting things {.smaller}

::: notes
we just went into groups how many groups are possible?
:::

In a class of 85 students how many groups of 4 are possible?

::: incremental
-   $85 \cdot 84 \cdot 83 \cdot 82$

-   $48,594,840$ groups

-   If we don't care about order

-   $\frac{85 \cdot 84 \cdot 83 \cdot 82}{4 \cdot 3 \cdot 2 \cdot 1}$

-   $2,024,785$ groups

-   *Combination* written as ${}_{n}C_{k}$ here $n = 85$, $k = 4$
:::

```{r}
#| echo: true
#| eval: true

exp(lfactorial(85))/exp(lfactorial(81))


exp(lfactorial(85))/(exp(lfactorial(81))*exp(lfactorial(4)))
```

## Examples

::: incremental
1. Let $Y \sim Bin(n = 5, p = 0.2)$ find the $E(Y)$ and $Var(Y)$
  a. $E(Y) = np = 1$
  b. $Var(Y) = np(1-p) = 0.8$
  c. What is the probability that $y = 3$?
     In `R: dbinom(size  = 5, prob = 0.2, x = 3)`
     $0.0512$
:::

## Warnings

10% Condition:  if trials are selected at random, it is OK to ignore dependence caused by sampling from a finite population  if the selected trials make up less than 10% of the population


# Poisson distribution - RIPS

::: incremental
-  Poisson - [fish](http://pecheur.info/en/gollum-fish-song.html) - Rips
- R - randomly through space or time
- I - indepedent
- P - proportional to interval size
- S - singly - no multiple occurences in space or time
:::

# Your turn

[Click here](https://stats1010-f22.github.io/website/ae/ae-8.html) or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-9.html")
plot(qr)
```
