---
title: "Lec 3 - Exploring categorical data"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
```

# Learning outcomes

By the end of this lesson you should:

-   Be able to identify categorical variables and why they are important

-   Graphical representation of two categorical variables in R

-   Tabular representation of two categorical variables

-   Graphical representation of one categorical variable

# Name some categorical variables

Use this [link](PollEv.com/gwynnsturdevant155), or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("PollEv.com/gwynnsturdevant155")
plot(qr)
```

## Definition of categorical variable

## The purpose of exploring categorical variables

-   Exploratory Data Analysis is about learning the structure of a dataset through a series of numerical and graphical techniques.

-   When you do EDA, you\'ll look for both

    -   general trends and

    -   interesting outliers in your data.

-   generate questions that will help inform subsequent analysis.
