---
title: "Chapter 12: The normal probability model"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
library(pdftools)
library(tidyverse)
```

# Learning outcomes

::: incremental
- The shape of the normal distribution and why it is important
- The central limit theorem.
:::

# {background-iframe="https://nzgwynn.shinyapps.io/limit_bin_normal/"}

## Binomial -> Normal

As the number of trials increases, the binomial pdf becomes well approximated by a normal distribution.

"Observed data often represent the accumulation of many small factors."

# Normal

$Y \sim N(\mu = 4, \sigma^2 = 25)$

::: incremental
- We use two parameters to describe a Normal distribution
- $Y \sim N(\mu = 0, \sigma^2 = 1)$ 
- $Y$ above is the *standard normal distribution*
:::

## Central limit theorem

::: notes
There are multiple versions of the CLT
:::

The probability distribution of a sum of independent random variables of comparable variance approaches a normal distribution as the number of summed random variables increases.

#  {background-iframe="https://rodrigojpereira.shinyapps.io/mstdteacher/"}

## Shifts & scales


::: columns
::: {.column width="50%"}
```{r shift}
#| echo: false
#| eval: true


x <- seq(-5, 15)
std_norm <- dnorm(x, mean = 0, sd = 1)
shifted_norm <- dnorm(x, mean = 10, sd = 1)


d <- tibble(x, std_norm, shifted_norm)

d %>% 
  ggplot() +
  geom_point(aes(x = x, y = std_norm)) +
  geom_point(aes(x = x, y = shifted_norm))

```
:::

::: {.column width="50%"}
```{r scale}
#| echo: false
#| eval: true


x <- seq(-15, 20)
std_norm <- dnorm(x, mean = 0, sd = 1)
scaled_norm <- dnorm(x, mean = 0, sd = 5)

d <- tibble(x, std_norm, scaled_norm)

d %>% 
  ggplot() +
  geom_point(aes(x = x, y = std_norm)) +
  geom_point(aes(x = x, y = scaled_norm))

```
:::
:::







# Your turn

[Click here](https://stats1010-f22.github.io/website/ae/ae-8.html) or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-9.html")
plot(qr)
```
