---
title: "Lec 4 - Exploring numeric data"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
```

# Learning outcomes

By the end of this lesson you should:

-   Understand measures of center and know how to compute them

-   Understand measures of spread and know how to compute them

-   Graphical representation and interpretation of one numerical variable in R

## Definition of numeric variable

A numeric or quantitative variable is a variable that can be measured.

## The purpose of Exploratory Data Analysis (EDA)

-   EDA is about learning the structure of a dataset through a series of numerical and graphical techniques.

-   When you do EDA, you'll look for both

    -   general trends and

    -   interesting outliers in your data.

generate questions that will help inform subsequent analysis.

# Observations {background-image="lect_04_images/observations.png" background-size="contain"}

# Variables {background-image="lect_04_images/variables.png" background-size="contain"}

# Measures of center

-   Mean

-   Median ($50\%th$ percentile)

# Mean

$\overline{var} =\frac{the\; sum\; of\; all\; the\; observations\; in\; the \; var}{the \; number \;of \;observations}$

# Mean vs median
```{r}
#| echo: false
#| eval: true

library(tidyverse)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 250) +
  geom_vline(aes(xintercept = mean(price)), color = "red") +
  geom_vline(aes(xintercept = median(price)), color = "green")

```
# Mean vs median

Mean - sensitive to outliers and skewed distributions

Median - more stable estimate 

Mode - most common value

# Measures of spread

-   Percentiles

    -   quartiles

    -   Interquartile range

- variance $s^2$
- standard deviation $s$
- range

# Variance

$s_{var}^2 =\frac{(observation\; 1 \; in\; the \; var - \overline{var})^2 + (observation\; 2 \; in\; the \; var - \overline{var})^2 + \dots + (observation\; n \; in\; the \; var - \overline{var})^2}{the \; number \;of \;observations - 1}$

# Graphically
```{r}
#| echo: false
#| eval: true

library(tidyverse)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 250) +
  geom_vline(aes(xintercept = mean(price)), color = "red")
  
```

# A few of the "squares"

```{r}
#| echo: false
#| eval: true

x_min <- c(2723, rep(3932.8, 4))
x_max <- c(3932.8, 5805, 7821, 5204, 5645)
y_min <- rep(0, 5)
y_max <- x_max - x_min

df <- data.frame(x_min, x_max, y_min, y_max)


ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 250) +
  geom_vline(aes(xintercept = mean(price)), color = "grey10") +
  geom_rect(data = df,
            aes(xmin = x_min, xmax = x_max, ymin = y_min, ymax = y_max), 
            colour = "lightblue", fill = "lightblue", alpha = 0.5)
```


# Standard deviation

$s_{var} =\sqrt{\frac{(observation\; 1 \; in\; the \; var - \overline{var})^2 + (observation\; 2 \; in\; the \; var - \overline{var})^2 + \dots + (observation\; n \; in\; the \; var - \overline{var})^2}{the \; number \;of \;observations - 1}}$

# Words to know

- mean
- median or quartile 2 (Q2)
- lower quartile (LQ) or quartile 1 (Q1)
- upper quartile (UQ) or quartile 3 (Q3)
- interquartile range (IQR)


# Your turn
[Click here](https://stats1010-f22.github.io/website/ae/ae-3.html) or the qr code below to write your first line of code

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-3.htm")
plot(qr)
```



