---
title: "Lec 5 - More probability rules"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
library(pdftools)
```

# Learning outcomes

By the end of this lesson you should:

-   Know and understand words like compliment, independent, and the Law of Large Numbers

-   Understand Boole's inequality

-   Understand the multiplication and addition rules

#  {background-image="lect_06-images/one_event.png" background-size="contain"}

# Law of large numbers ($\infty$ in practice)
```{r}
flips <- sample(c("heads", "tails"), size = 500, replace = TRUE)
plot(cumsum(flips == "heads") / (1:length(flips)), 
     type = "l", ylim = c(0,1),  
     main = "Coin Flips", 
     xlab = "Number of flips", ylab = "Proportion of heads")
abline(h = 0.5, col = "red")
```

# Compliment of $A$

$S - A$

#  {background-image="lect_06-images/blue_zones_venn_diagram.png" background-size="contain"}

# Addition rule

$Pr(A \cup B)=Pr(A)+Pr(B)−Pr(A \cap B)$

# Boole's inequality

$Pr(A)+Pr(B) \leq Pr(A \cup B)$

# Independent events

Events are said to be independent if the outcome of one event does not impact the probability of another.

Which of these events is independent?

\- Removing balls from urns *with replacement*

\- Removing balls from urns *without replacement*

\- Flipping a coin ten times

\- Riding in an uber and getting a free meal

\- Going into a store and purchasing something from that store

\- Suitability of one two employees for a job

\- Drawing 2 kings from a deck of cards in a row *without replacement*

# 3 mins

```{r}
countdown(minutes = 3, seconds = 0, style = "position: relative; width: min-content;")
```

# Conditional probability

When events are not independent conditional probabilities are useful.

We use the notation: $$Pr(Card\: 2\: is\: a\: king∣Card\: 1\: is\: a\: king)$$ for this. We use the ∣ as shorthand for "given that" or "conditional on".

# Multiplication rule

$$Pr(A \cap B) = Pr(A)Pr(B∣A)$$

when $A$ and $B$ are independent, $Pr(B|A)$ = ?

::: incremental
-   $Pr(B|A) = Pr(B)$

-   $Pr(A \cap B) = Pr(A)Pr(B)$
:::

# Tree diagrams



# Your turn

[Click here](https://stats1010-f22.github.io/website/ae/ae-5.html) or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-5.html")
plot(qr)
```
