---
title: "Chapter 12: The normal probability model"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
library(pdftools)
library(tidyverse)
```

# Learning outcomes

::: incremental
- The shape of the normal distribution and why it is important
- The central limit theorem.
:::

# {background-iframe="https://nzgwynn.shinyapps.io/limit_bin_normal/"}

## Binomial -> Normal

As the number of trials increases, the binomial pdf becomes well approximated by a normal distribution.

"Observed data often represent the accumulation of many small factors."

# Normal

$Y \sim N(\mu = 4, \sigma^2 = 25)$

::: incremental
- We use two parameters to describe a Normal distribution
- $Y \sim N(\mu = 0, \sigma^2 = 1)$ 
- $Y$ above is the *standard normal distribution*
:::

## Central limit theorem

::: notes
There are multiple versions of the CLT
:::

The probability distribution of a sum of independent random variables of comparable variance approaches a normal distribution as the number of summed random variables increases.

#  {background-iframe="https://rodrigojpereira.shinyapps.io/mstdteacher/"}

## Shifts & scales

::: columns
::: {.column width="50%"}
```{r shift}
#| echo: false
#| eval: true


x <- seq(-5, 15)
std_norm <- dnorm(x, mean = 0, sd = 1)
shifted_norm <- dnorm(x, mean = 10, sd = 1)


d <- tibble(x, std_norm, shifted_norm)

d %>% 
  ggplot() +
  geom_point(aes(x = x, y = std_norm)) +
  geom_point(aes(x = x, y = shifted_norm))

```
:::

::: {.column width="50%"}
```{r scale}
#| echo: false
#| eval: true


x <- seq(-15, 20)
std_norm <- dnorm(x, mean = 0, sd = 1)
scaled_norm <- dnorm(x, mean = 0, sd = 5)

d <- tibble(x, std_norm, scaled_norm)

d %>% 
  ggplot() +
  geom_point(aes(x = x, y = std_norm)) +
  geom_point(aes(x = x, y = scaled_norm))
```
:::
:::

## Standardizing {.smaller}

::: notes
Books and tables of just normal probabilities to multiple decimal places.
:::

Historically, it could be quite hard to find probabilities, so *standardizing* was important.

::: incremental
- use shift and scale information from above
- $Z = \frac{X-\mu_X}{\sigma_X}$
- Find $E(Z)$ and $Var(Z)$
- $E(Z) = E(\frac{X-\mu_X}{\sigma_X})$
- $$\begin{aligned}
E(Z) & = E(\frac{X-\mu_X}{\sigma_X})\\ 
 & = \frac{1}{\sigma}E(X-\mu_X) \\
 &= 0
 \end{aligned}$$
- $$\begin{aligned}
Var(Z) & = Var(\frac{X-\mu_X}{\sigma_X})\\ 
 & = \frac{1}{\sigma^2}Var(X-\mu_X) \\
 &= \frac{\sigma^2}{\sigma^2} = 1
 \end{aligned}$$
:::


## Example

Let $Y \sim N(\mu_Y = 5, \sigma_Y^2 = 4)$, standardize the following and find the probability:

::: incremental
1. $P(Y < 3)$
- $$\begin{aligned}
P(Y < 7) &= P(\frac{Y - \mu_Y}{\sigma_Y} < \frac{3-\mu_Y}{\sigma_Y}) \\ 
 & = P(Z < \frac{3-5}{2} = -1)
 \end{aligned}$$
 - `pnorm(-1)` $0.1586553$
 - `pnorm(-1, mean = 5, sd = 2)`
:::



#  {background-image="lec_11-things/p_less_neg_1.png" background-size="contain"}

## Queen Bee `Peenorm`

::: notes
To the left, to the left, pnorm is to the left
:::

{{< video https://www.youtube.com/embed/Nw8LyN21OO8 >}}


## Example

::: notes
Both of these probabilities are the same because of *symmetry* of the normal distribution. How does this generalize to all normal distributions?
:::

Let $Y \sim N(\mu_Y = 5, \sigma_Y^2 = 4)$, standardize the following and find the probability:

::: incremental
2. $P(Y > 7)$
- $$\begin{aligned}
P(Y > 7) &= P(\frac{Y - \mu_Y}{\sigma_Y} > \frac{7-\mu_Y}{\sigma_Y}) \\ 
 & = P(Z > \frac{7-5}{2} = 1)
 \end{aligned}$$
- `1- pnorm(1)` $ 0.1586553$ 
- `1- pnorm(1, mean = 5, sigma = 2)`
:::

#  {background-image="lec_11-things/p_greater_1.png" background-size="contain"}


#  {background-image="lec_11-things/emperical_rule.png" background-size="contain"}


#  {background-image="lec_11-things/q_norm.png" background-size="contain"}

# Percentiles
::: notes
Earlier we used the normal to find a percent, you can also use `R` to find the z value for a certain percent.
:::

::: incremental

1. Find $z : P(Z < z) = .1711$
- `qnorm(.1711)`
2. A medical test produces a score that measures the risk of a disease. In healthy adults, the test score is normally distributed with $\mu=10$ point and $\sigma = 2.5$. Lower scores suggest the disease is present. Test scores below what threshold signal a problem for only $1\%$ of healthy adults?
- `qnorm(.01, mean = 10, sd = 2.5)`

:::

# Normality tests


# Your turn

[Click here](https://stats1010-f22.github.io/website/ae/ae-8.html) or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-9.html")
plot(qr)
```
