---
title: "Chapter 6: Association between quantitative variables"
subtitle: "STAT 1010 - Fall 2022"
footer:  "[stat1010-f22.github.io/website](https://stats1010-f22.github.io/website/)"
format: 
  revealjs: 
    theme: slides.scss
    transition: fade
    slide-number: true
editor: visual
execute:
  freeze: auto
---

```{r include = FALSE}
library(countdown)
library(pdftools)
library(tidyverse)
```

# Learning outcomes

By the end of this lesson you should:

-   Know how to read a scatterplot to describe associations between quantitative variables

-   

# Revision - associations

::: {.notes}
Remind them of $\chi^2$ tests
:::

::: {.incremental}
- $\chi^2$ tests are used for associations in categorical data
- What is used to find associations in numerical data?
:::

# Plots
::: {.notes}
Remind them about plots of proportions that we did for titanic and comics data.

Which variable goes on the y axis?

Is there an association or not?
:::

::: {.incremental}
- variable on the $y$ axis 
    - dependent
    - response
- variable on the $x$ axis
    - independent
    - explanatory
    
```{r scatterplot}

diamonds %>% 
  filter(x > 0) %>% 
  ggplot() +
  geom_point(aes(y = price, x = x)) 
```
:::

# Visual test

::: columns
::: {.column width="50%"}
```{r pdf-add-run}
#| eval: true
#| message: false
#| echo: false
#| column: page-left

diamonds %>% 
  filter(x > 0) %>% 
  ggplot() +
  geom_point(aes(y = price, x = x)) 
  
  
```
:::

::: {.column width="50%"}
```{r pdf-no_add}
#| eval: true
#| message: false
#| echo: false
#| column: page-right

diamonds %>% 
  filter(x > 0) %>% 
  ggplot() +
  geom_point(aes(y = price, x = x)) 

```
:::
:::



# Your turn

[Click here](https://stats1010-f22.github.io/website/ae/ae-8.html) or the qr code below

```{r, eval=TRUE, echo=FALSE}
library(qrcode)
qr <- qr_code("https://stats1010-f22.github.io/website/ae/ae-9.html")
plot(qr)
```
