---
title: "Introduction to the diamonds dataset"
output: 
   html_document
---

::: callout-important
This application exercise is due at noon on 6 Sept at 11:59am.
:::

# Scenario

You own a jewelry business and are working to price your [diamonds](tps://www.londonde.com/blogs/worlds-most-famous-diamonds/). To facilitate, you **explore** the diamond dataset in R and pay special attention to the cost of the diamond. Let's **explore** together.

# Coding basics

[Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread.](https://style.tidyverse.org/index.html)

-   When writing code, load all libraries first.

-   "\##" is used to write comments in R and should be used to justify next steps and anything unusual or unexpected steps

-   Comments are really important and are notes to your future self to remind you about steps that you took and decisions that you made.

```{r load-packages, eval=FALSE, echo=TRUE}
#| warning: false
#| message: false

## loading library
library(tidyverse) # for data analysis and visualisation
library(plotly) # for interactive plots
```

# Data

The `diamonds` dataset includes basic information about 53940 diamonds including price and other characteristics.

```{r read-data, echo=TRUE, eval=FALSE}
#| message: false
#| warning: false

## View the dataset in a separate tab
View(diamonds)
```

The \_\_\_ dataset has \_\_\_ observations and \_\_\_ variables.

What are these variables and what do they mean?

```{r help-on-dataset, eval=FALSE, echo=TRUE}
## What are the variables in this dataset?
## What are they counting?
?diamonds
```

1.  What do the "4 c's" of every diamond mean? What kind of variables are they? (hint: This [video](https://www.youtube.com/watch?v=FqB5Es1HXI4) may help, and please copy this into your notes  under the title data types) 

```{r plot-price, eval=FALSE, echo=TRUE}
#| eval: false
#| echo: true
## What is the distribution of price?
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 1) 
```

2. What kind of variable is price? What happens if you draw a histogram with another kind of variable?

```{r plot-other, eval=FALSE, echo=TRUE}
#| eval: false
#| echo: true

## What is the distribution of price?
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = OTHER_VAR), binwidth = 1) 
```

3.  What does the binwidth in the geom_histogram() function do? (hint: draw multiple plots changing this value) What binwidth value is most appropriate for this data?

```{r plot-other, eval=FALSE, echo=TRUE}
#| eval: false
#| echo: true

## What is the distribution of price?
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = OTHER_VALUE) 
```

The distribution of diamond prices is right skewed (long right tail).

4.  In your notes, copy [these histograms](https://www.statology.org/describe-shape-of-histogram/) and the words used to describe them.

One interesting feature of this graph is the dip in diamonds priced at about \$1000. Let's explore this a little further and look at some basic [dplyr commands](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf).

```{r, diamonds_about_1000_a, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true

diamonds %>% 
  filter(price > 750 & price < 1250) %>% 
  View()
```

5.  What does the filter function do?

```{r diamonds_about_1000_b, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true

diamonds %>% 
  filter(price > 750 & price < 1250) %>% 
  mutate(price_hundred = price %/% 100) %>% 
  View()
```

6.  What does this "%/%" operator do? (hint: run 10 %/% 3, 10 %/% 5, 7 %/% 3, 7 %/% 5 in the console)

7. What does the mutate function do? (hint: pay special attention to the dimensions of the data)

```{r diamonds_about_1000_c, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true

diamonds %>% 
  filter(price > 750 & price < 1250) %>% 
  mutate(price_hundred = price %/% 100) %>% 
  select(price, price_hundred) %>% 
  View()
```

8. What does the select function do?

```{r diamonds_about_1000_d, eval=FALSE, echo=TRUE}
#| eval: false
#| echo: true


diamonds %>% 
  filter(price > 750 & price < 1250) %>% 
  mutate(price_hundred = price %/% 100) %>% 
  select(price, price_hundred) %>% 
  count(price_hundred) %>% 
  View()
```

9. Can you identify the values where the dip is using the information above? Why or why not?

```{r diamonds_about_1000_e, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true

diamonds %>% 
  filter(price > 750 & price < 1250) %>% 
  mutate(price_hundred = price %/% 100) %>% 
  select(price, price_hundred) %>% 
  mutate(price_remainder = price %% 100) %>% 
  count(price_hundred, price_remainder) %>% 
  View()
```

10  What does the "%%" function do?

11. Can you identify the values where the dip is using the information above? Why or why not? We will address this later in the semester.


Cut may impact upon price

```{r cut_plot, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, color = cut, fill = cut),
                 binwidth = 100) 
```

12.  Which cut do you suspect is the most common in the dataset? How many diamonds with that cut are in the dataset?

```{r cut_count, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true
#| 
diamonds %>% 
  INSERT_FUNCTION_HERE(cut)
```

13.  The variables color or clarity might impact upon the price. Draw a histogram and color it using one of them.

```{r cut, eval=FALSE, echo=TRUE}

#| eval: false
#| echo: true
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, color = VAR_NAME, 
                               fill = VAR_NAME), 
                 binwidth = YOUR_BIN_WIDTH) 
```

14. What do the fill and color inputs do in the above coding?

15. What kind of variables should we use for the fill and color inputs? What happens if another type of variable is used?

```{r other_vars, echo=TRUE, eval=FALSE}
#| eval: false
#| echo: true

ggplot(data = diamonds) +
      geom_histogram(mapping = aes(x = price, color = VAR_NAME, 
                                   fill = VAR_NAME), 
                     binwidth = YOUR_BIN_WIDTH) 
```

Statisticians use basic summaries to desribe variables. Some of them are mean, median, and mode. 


