---
title: "Revision 1 for exam 1"
output: 
   html_document
---

```{r}
#| eval: true
#| message: false
library(tidyverse)
```


1. What tidyverse function would you use to extract rows?
  a. filter()
  b. count()
  c. select()
  d. distinct()


2. What tidyverse function would you use to extract colums?
  a. filter()
  b. count()
  c. select()
  d. distinct()

3. To find the maximum number of times that a categorical variable appears which function should I use?
  a. filter()
  b. count()
  c. select()
  d. distinct()

4. What does the "%>%" operator do?
  a. used in tidyverse between layers of dplyr coding
  b. used in ggplot to add more layers of coding

5. Will this code run without an error? 
mutate(price_hundred = price %/% 100)
  a. It will produce an error
  b. It will not produce an error

6. What does the binwidth input do?
  a. It controls the 
  

7. What punctuation must be placed before a function or dataset to access more information?
  a. "."
  b. "+"
  c. "?"
  d. "!"

9. Where should the variable names go in the following line of code: 
ggplot(data = <A>) + 
  <B>(mapping = aes(<C>))
  a. A
  b. B
  c. C
  

10. What does the here package do?
  a. for data manipulation and display
  b. to organize files
  c. none of the above


11. Which of the following variables are discrete?
 a. number of employees in a company.
 b. distance 



15. What is the difference between distinct() and count()?
a. they are the same
b. distinct() gives the levels in a variable, count() gives the marginal distribution
c. distinct() gives the marginal distribution, count() gives the levels in a variable

16. What operator is used in ggplot? 
  a. "."
  b. "+"
  c. "?"
  d. "!"

Categorical variables

15. Find the expected count for clarity VS1 and color I for a $\chi^2$ distribution assuming independence.

```{r chi-squared}
#| eval: true
#| echo: false

diamonds %>% 
  count(color, clarity) %>% 
  pivot_wider(names_from = color,
              values_from = n) %>% 
  mutate(marginal_clarity = rowSums(.[2:7])) %>% 
  mutate(marginal_color = colSums(.[2:9]))
  
```



16. Some of the last questions on these - https://openintro.shinyapps.io/ims-02-explore-01/#section-distribution-of-one-variable


17. What proportion of the very good cut diamonds are colored I?

```{r prop_color}
#| eval: true
#| echo: false

diamonds %>% 
  count(color, cut) %>% 
  group_by(color) %>% 
  mutate(prop = n/ sum(n)) %>% 
  select(color, cut, prop) %>% 
  pivot_wider(names_from = color,
              values_from = prop) 
  
```
```{r prop_cut}
#| eval: true
#| echo: false

diamonds %>% 
  count(color, cut) %>% 
  group_by(cut) %>% 
  mutate(prop = n/ sum(n)) %>% 
  select(color, cut, prop) %>% 
  pivot_wider(names_from = color,
              values_from = prop) 
  
```

19. Which of the lines below is  more likely to be the mean and which is likely to be the median, and which is likely to be the mode.
```{r, mean-median}
#| echo: false
#| eval: true

library(tidyverse)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 250) +
  geom_vline(aes(xintercept = mean(price)), color = "red") +
  geom_vline(aes(xintercept = median(price)), color = "green")

```
a. red is median, green is mean
b. red is mean, green is mode
c. red is mode, green is mean
d. red is mean, green is median

20. Is the mean or median a better measure of center in this distribution.
```{r, mean-median}
#| echo: false
#| eval: true
#| warning: false

library(tidyverse)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = x), binwidth = .4) +
  xlim(2, 10)
```

21. Which plot is better to display a outliers:
a. histogram
b. dotplot
c. plot of proportions
d. boxplot

