---
title: "Revision 1 for exam 1"
output: 
   html_document
---

```{r}
#| eval: true
#| message: false
#| echo: false
library(tidyverse)
```


1. What tidyverse function would you use to extract rows?
  *a. filter()*
  b. count()
  c. select()
  d. distinct()


2. What tidyverse function would you use to extract colums?
  a. filter()
  b. count()
  *c. select()*
  d. distinct()

3. To find the maximum number of times that a categorical variable appears which function should I use?
  a. filter()
  *b. count()*
  c. select()
  d. distinct()

4. What does the "%>%" operator do?
  *a. used in tidyverse between layers of dplyr coding*
  b. used in ggplot to add more layers of coding

5. Will this code run without an error? 
diamonds %>% 
   mutate(price_hundred = price %/% 100)
  a. It will produce an error
  *b. It will not produce an error*

6. What does the binwidth input in ggplot do?
  a. It controls the opacity of the plot
  b. It can be used to update the labels
  *c. It controls the smoothness of a density plot*
  d. It controls the size of dots
  

7. What punctuation must be placed before a function or dataset to access more information?
  a. "."
  b. "+"
  *c. "?"*
  d. "!"

9. Where should the variable names go in the following line of code: 
ggplot(data = A) + 
  B(mapping = aes(C))
  a. A
  b. B
  *c. C*
  

10. What does the here package do?
  a. for data manipulation and display
  *b. to organize files*
  c. none of the above


11. Which of the following variables are discrete?
 *a. number of employees in a company.*
 b. distance traveled to and from work
 c. taxes paid in 2019
 d. purchases from 2018
 


15. What is the difference between distinct() and count()?
a. they are the same
*b. distinct() gives the levels in a variable, count() gives the marginal distribution*
c. distinct() gives the marginal distribution, count() gives the levels in a variable

17. What operator is used in ggplot?
  a. "."
  *b. "+"*
  c. "?"
  d. "!"

Categorical variables

18. Find the expected count for clarity VS1 and color I for a $\chi^2$ distribution assuming independence.



```{r chi-squared}
#| eval: true
#| echo: false

diamonds %>% 
  count(color, clarity) %>% 
  pivot_wider(names_from = color,
              values_from = n) %>% 
  mutate(marginal_clarity = rowSums(.[2:8])) %>% 
  mutate(marginal_color = colSums(.[2:9]))

8171*2808/53940
```

19. What proportion of the very good cut diamonds are colored I?
*0.0997*

```{r prop_color}
#| eval: true
#| echo: false

diamonds %>% 
  count(color, cut) %>% 
  group_by(color) %>% 
  mutate(prop = n/ sum(n)) %>% 
  select(color, cut, prop) %>% 
  pivot_wider(names_from = color,
              values_from = prop) 
  
```
```{r prop_cut}
#| eval: true
#| echo: false

diamonds %>% 
  count(color, cut) %>% 
  group_by(cut) %>% 
  mutate(prop = n/sum(n)) %>% 
  select(color, cut, prop) %>% 
  pivot_wider(names_from = color,
              values_from = prop) 
  
```

20. Which of the lines below is the mean, median, or mode?
```{r, mean-median}
#| echo: false
#| eval: true

library(tidyverse)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 250) +
  geom_vline(aes(xintercept = mean(price)), color = "red") +
  geom_vline(aes(xintercept = median(price)), color = "green")

```
a. red is median, green is mean
b. red is mean, green is mode
c. red is mode, green is mean
*d. red is mean, green is median*

21. Is the mean or median a better measure of center in this distribution.

*The mean is a better estimate in a a symmetric distribution, median in a skewed distribution*
```{r, hist-x}
#| echo: false
#| eval: true
#| warning: false

library(tidyverse)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = x), binwidth = .4) +
  xlim(2, 10)
```

22. Which plot is best at displaying outliers:
a. histogram
b. dotplot
c. plot of proportions
*d. boxplot*

