---
title: "Random variables"
output: 
   html_document
---

::: callout-important
This application exercise is due on 21 Oct at 2:00pm.
:::

```{r}
#| eval: false
#| message: false
#| echo: true
library(tidyverse) # for data manipulation and plots
library(praise) # for good vibes
```

Chapter 6: Association between quantitative variables

1.  If the covariance between $x$ and $y$ is $0$, then the correlation between $x$ and $y$ is $0$ as well

    a.  True
    b.  False

2.  A retailer calculated the correlation line between the price of an item ($x$) and the amount sold ($y$). The correlation line is the same if the $x$ and $y$ variables are exchanged.

    a.  True
    b.  False
    

```{r}
#| eval: false
#| message: false
#| echo: true

praise()
```

3.  If the correlation between number of customers and sales in dollars in retail stores is $r=0.6$, then what would be the correlation if the sales were measured in thousands of dollars? In euros? (1 euro is worth about $\$US0.97$)


