---
title: "Introduction to the diamonds dataset"
editor: visual
---

# Scenario

You own a jewelry business and are working to price your [diamonds](tps://www.londonde.com/blogs/worlds-most-famous-diamonds/). To facilitate, you **explore** the diamond dataset in R and pay special attention to the cost of the diamond. Let's **explore** together.

# Coding basics

[Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread.](https://style.tidyverse.org/index.html)

-   When writing code, load all libraries first.

-   "\##" is used to write comments in R and should be used to justify next steps and anything unusual or unexpected steps

-   Comments are really important and are notes to your future self to remind you about steps that you took and decisions that you made.

```{r, eval=FALSE, echo=TRUE}
## loading library
library(tidyverse)

## First look at the dataset
View(diamonds)
```

```{r, eval=FALSE, echo=TRUE}
## What is in this dataset?
?diamonds
```

```{r, eval=FALSE, echo=TRUE}
## What is the distribution of price?
ggplot(data = diamonds, aes(x = price)) +
  geom_dotplot(stackratio = 0.009, alpha = 0.05, stackdir = "centerwhole") 
```

## What should the price of your diamonds be?

Based upon the above plot, and since the diamonds are nice, they should all cost above $150,000, right?


